---
title: "Access taxon or community occurrences"
output: 
  rmarkdown::html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: yeti
    highlight: tango
vignette: >
  %\VignetteIndexEntry{occurrences}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(tscr)
```

# Setup tscr
Read how to [setup](https://dbca-wa.github.io/tscr/articles/setup.html) tscr
to the TSC API.

# Download Occurrences from TSC
We get Taxon occurrences with their point representation from TSC.
Currently, all Fauna and Flora occurrences have only points, not polygons.
Flora Populations can be captured as polygons representing the entire
population, or a surveyed part of it.
While some taxon occurrences might have polygon extents, all have a point 
representation.

Community occurrences are mostly captured as polygons. Each community polygon 
has a corresponding point representation (the centroid), but we limit this example
to their polygons.

The examples are mutually applicable to taxon polygons and community points.

## Areas as table
TSC maintains several types of reference areas, such as DBCA Regions and Districts. 

In this example, we download a subset of 10 Areas, parse the data into a table,
and display selected columns of the downloaded Areas.

```{r areas}
areas <- tsc_GET("area", chunk_size = 100, max_records = 10) %>% tsc_parse()

areas %>% 
  dplyr::select("area_type", "name") %>% 
  dplyr::arrange("area_type") %>% 
  janitor::clean_names(case="sentence") %>% 
  knitr::kable()
```

## Areas on a map

In this example, we download Areas of all types, convert it to a spatial format
(GeoJSON), and display it on a map. 

See the [Areas API](https://tsc.dbca.wa.gov.au/api/1/area/) for other possible
filters.

We set `mapview::mapviewOptions` with some sensible defaults.
See the [mapview](https://r-spatial.github.io/mapview/index.html) homepage for 
far more map configuration options.

```{r areas_gj, eval=T}
areas_gj <- 
  tsc_GET("area", 
          # query = list(area_type="Region"),
          chunk_size = 100) %>% 
  magrittr::extract2("data") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf()

mapview::mapviewOptions(basemaps = c("Esri.WorldImagery", 
                                     "Esri.WorldShadedRelief", 
                                     "OpenTopoMap",
                                     "OpenStreetMap"),
                        layers.control.pos = "topright")

mapview::mapview(areas_gj, zcol="area_type")
```

## Download Occurrences
In this example, we download the point representation of Taxon occurrence records.

We limit the result size to 10 records for demonstration purposes, but this
method can also download the entire dataset if the parameter `max_records` is
omitted.

See the [TSC API](https://tsc.dbca.wa.gov.au/api/1/) for the full list of available
endpoints. 

We show here 
[Taxon Point Occurrences](https://tsc.dbca.wa.gov.au/api/1/occ-taxon-points/)
and [Community Polygon Occurrences](https://tsc.dbca.wa.gov.au/api/1/occ-community-areas/).

Each data field can serve as a filter via the query attribute.


```{r get_tae}
# Taxon occurrences with their Point representation
tae_res <- tsc_GET("occ-taxon-points", max_records = 10)

# Community occurrences with their Polygon representation
cae_res <- tsc_GET("occ-community-areas", max_records = 10)
```

## Occurrences as table
We show some fields as an interactive table. The field `as_html` contains 
rendered content which can be used as a map popup.

### Taxon points
```{r tae_tbl}
tae <- tae_res %>% tsc_parse()
# head(tae) %>% knitr::kable()
tae %>% 
  dplyr::select(label, encountered_on, geolocation_capture_method, as_html) %>% 
  janitor::clean_names(case="sentence") %>% 
  reactable::reactable(filterable=TRUE, sortable = TRUE, searchable = TRUE,
                      defaultColDef = reactable::colDef(html=TRUE))
```


### Community areas
```{r cae_tbl}
cae <- cae_res %>% tsc_parse()
cae %>% 
  dplyr::select(label, encountered_on, as_html) %>% 
  janitor::clean_names(case="sentence") %>% 
  reactable::reactable(filterable=TRUE, sortable = TRUE, searchable = TRUE,
                      defaultColDef = reactable::colDef(html=TRUE))
```

## Occurrences on a map
Taxon occurrences as GeoJSON SimpleFeatures for plotting.
We use the field `as_html` as the popup and the geolocation capture method for
grouping.

### Taxon map
```{r tae_map}
tae_gj <- tae_res %>%
  magrittr::extract2("data") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf()

mapview::mapview(tae_gj, popup="as_html", zcol="geolocation_capture_method")
```

### Community map
```{r cae_map}
cae_gj <- cae_res %>%
  magrittr::extract2("data") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf()

mapview::mapview(cae_gj, popup="as_html", zcol="name")
```

# Download all Occurrences from TSC
We get Taxon occurrences with their point representations from TSC through the 
API endpoint [occ-taxon-points](https://tsc.dbca.wa.gov.au/api/1/occ-taxon-points/).

Currently, all Fauna and Flora occurrences are georeferenced only with point 
coordinates, not polygons.

We then get TSC Areas and split them into DBCA Regions and Districts.
Caveat: Regions and Districts to not comprehensively cover WA.

Lastly, we get a list of taxonomic names from TSC to resolve taxon IDs from
taxon occurrences to names.

Taxon Occurrences (taxon points) are presented as `occ` with DBCA Region and District 
names (where applicable), plus taxonomic names.
Community Occurrences (community areas) are named `com_occ`. 

```{r download, eval=F}
tsc_tae <- tsc_GET("occ-taxon-points") # 137k records
tsc_cae <- tsc_GET("occ-community-areas")
tsc_taxa <- tsc_GET("taxon") %>%
    wastdr::wastd_parse() %>%
    dplyr::mutate(name_id = name_id %>% as.character(), taxon = pk)
tsc_areas <- tsc_GET("area") %>%
      magrittr::extract2("features") %>%
      geojsonio::as.json() %>%
      geojsonsf::geojson_sf()

save(tsc_tae, tsc_cae, tsc_taxa, tsc_areas, 
     file=here::here("data/tsc.RData"), compress="xz")
```

```{r munge, eval=F}
regions <- tsc_areas %>%
  dplyr::filter(area_type == "Region") %>%
  dplyr::transmute(region_id = pk, region_name = name)

districts <- tsc_areas %>%
  dplyr::filter(area_type == "District") %>%
  dplyr::transmute(district_id = pk, district_name = name)

occ <- tsc_tae %>%
  magrittr::extract2("features") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf() %>% 
  sf::st_join(regions) %>% 
  sf::st_join(districts) %>% 
  dplyr::left_join(tsc_taxa, by="taxon")

com_occ <- tsc_cae %>%
  magrittr::extract2("features") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf() %>% 
  sf::st_join(regions) %>% 
  sf::st_join(districts)
```

# Analyse occurrences

## Taxon occurrences by DBCA Region
This example demonstrates how to tally taxon occurences by DBCA Regions, which
were joined spatially earlier.
```{r occ_by_reg, eval=FALSE}
occ_by_region <- occ %>% 
    dplyr::group_by(region_name, taxonomic_name) %>% 
    dplyr::tally()  %>% 
    wastdr::sf_as_tbl() %>% 
    tidyr::pivot_wider(
        id_cols=taxonomic_name,
        names_from = region_name,
        values_from = n,
        values_fill = NA
    )
occ_by_region %>% reactable::reactable(sortable = T, filterable = T)
```

## Species occurrences by DBCA District
The same summary, but by District.
```{r occ_by_dist, eval=FALSE}
occ_by_district <- occ %>% 
    dplyr::group_by(district_name, taxonomic_name) %>% 
    dplyr::tally()  %>% 
    wastdr::sf_as_tbl() %>% 
    tidyr::pivot_wider(
        id_cols=taxonomic_name,
        names_from = district_name,
        values_from = n,
        values_fill = NA
    )
occ_by_district %>% reactable::reactable(sortable = T, filterable = T)
```
