---
title: "Setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tscr)
```

This vignette will show how to set up `tscr` via environment variables to connect
to either the TSC API, or the TSC UAT API (to test write operations on disposable data).

## Settings via environment variables
Run `usethis::edit_r_environ()` to open your `.Renviron` file.
Paste the following code into the file with your own TSC credentials, then
restart your R session (Ctrl+Shift+F10 in RStudio) to apply the changes.

`tscr` will use these environment variables as defaults for the TSC API location
and authentication.

```{r, eval=FALSE}
TODO
```

## Settings via setup
To jump between TSC instances, you can set up `tscr` via `tscr::tscr_setup()`.

```{r, eval=FALSE}
tscr::setup(
  api_url = "https://tsc.dbca.wa.gov.au/api/1/",
  api_token = "Token xxxx",
  verbose = TRUE
)
# typically followed by several functions requiring above settings
```

## Settings via parameters
You can run individual functions with specific settings as parameters.
The tests set the TSC UAT credentials explicitly this way.

```{r, eval=FALSE}
tscr::tsc_GET(
  serializer = "area",
  api_url = "https://tsc.dbca.wa.gov.au/api/1/",
  api_token = "Token xxxx",
  verbose = TRUE
)
```

