---
title: "Update taxonomic names from WACensus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{wacensus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Context

* WACensus, an Oracle database, is the point of truth for WA taxonomic names at DBCA.
* KMI publishes a few WACensus views as GeoServer layers.
* TSC maintains a copy of those WACensus views and rebuilds its internal taxonomy
  from this copy of WACensus data.

This workbook shows how to update TSC from KMI's WACensus views.

`gjprod` is defined here as a shortcut to `tscr::upsert_geojson` for convenience.

```{r setup}
library(tscr)
gjprod <- function(data, ser) upsert_geojson(
  data, serializer = ser, api_url = get_tsc_api_url(), chunksize = 1000)
```

## Import a copy of WACensus into TSC
Run against local, UAT or production TSC instance. 
Adjust chunksize down when getting "504 gateway timeouts".

```{r etl, eval=F}
"public:herbie_hbvsupra_public" %>% gs_getWFS() %>% gjprod("supra")
"public:herbie_hbvgroups_public" %>% gs_getWFS() %>% gjprod("groups")
"public:herbie_hbvnames_public" %>% gs_getWFS() %>% gjprod("names")
"public:herbie_hbvfamilies_public" %>% gs_getWFS() %>% gjprod("families")
"public:herbie_hbvgenera_public" %>% gs_getWFS() %>% gjprod("genera")
"public:herbie_hbvspecies_public" %>% gs_getWFS() %>% gjprod("species")
"public:herbie_hbvxrefs_public" %>% gs_getWFS() %>% gjprod("xrefs")
"public:herbie_hbtparents_public" %>% gs_getWFS() %>% gjprod("parents")
"public:herbie_hbvvernaculars_public" %>% gs_getWFS() %>% gjprod("vernaculars")
```

## Rebuild TSC taxonomy
TSC > Curators > Update Taxon Names. After a few minutes, TSC Taxa will be updated.


# Further examples
## Update one Taxon
This section demonstrates how to update a single taxon in TSC.

```{r get_single}
hbv_groups <- gs_getWFS(layer_name = "public:herbie_hbvgroups_public")
groups_props <- purrr::map(hbv_groups[["features"]], "properties")
groups_props[[1]]
```
This is the format of the data expected by `tsc_POST`.

```{r upload_single, eval=FALSE}
tsc_POST(groups_props[[1]], serializer = "group")
```

## List taxonomic names
Similarly, we could extract the list of taxonomic names from the KMS GeoServer
layer `public:herbie_hbvnames_public`, which represents the WACensus view `hbvnames`.

We retain the fields used in TSC and limit the results to 100 names to keep
the file size of this vignette small.

```{r list_names}
wace_names <- "public:herbie_hbvnames_public" %>% 
  gs_getWFS() %>% 
  magrittr::extract2("features") %>% 
  {tibble::tibble(
    name_id = purrr::map(., c("properties", "name_id")) %>% as.integer(),
    name = purrr::map(., c("properties", "name")) %>% as.character(),
    author = purrr::map(., c("properties", "author")) %>% as.character(),
    reference = purrr::map(., c("properties", "reference")) %>% as.character(),
    rank_name = purrr::map(., c("properties", "rank_name")) %>% as.character()
  )} %>% 
  head(100)

wace_names %>% 
  reactable::reactable(filterable = TRUE, sortable = TRUE, searchable = TRUE)
```